A 38 Million Year Old Mini-Neptune in the Kepler Field

### 2021-09-21 paper/ms.tex
update ms
Co-authored-by: Kento Masuda <kmasuda@ess.sci.osaka-u.ac.jp>

% 208 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously known to host a
  $3.3\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler satellite because it is nearby
  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Here we show
  using Gaia kinematics, TESS stellar rotation periods, and
  spectroscopic lithium abundances that Kepler\,1627 is a member of
  the \clusterage\ old $\delta$~Lyr cluster.  To our knowledge,
  Kepler\,1627Ab is the youngest planet with a precise age yet found
  by the main Kepler mission.  From unresolved starspot-crossing
  events in the Kepler photometry and a Keck/HIRES transit we find
  that the planet is likely on an orbit aligned with the stellar spin
  axis.  We also briefly explore the flare arrival time distribution
  in the short cadence Kepler photometry.  The $\delta$~Lyr cluster is
  one of many stellar groups whose properties have been clarified
  using Gaia; many other new clusters overlap with confirmed and
  candidate exoplanets.  Verifying the results of Gaia clustering
  analyses using TESS rotation periods offers a promising path toward
  expanding the census of age-dated planets, and our corresponding
  understanding of planetary evolution.
\end{abstract}

@@

The K2 and TESS missions have now enabled the detection of about ten
close-in planets younger than 100 million years, all smaller than
Jupiter
\citep{Mann_K2_33b_2016,David_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,plavchan_planet_2020,rizzuto_tess_2020,martioli_aumicbc_2021}.
The Kepler mission however has not yielded any planets with precise
ages below one gigayear \citep{Meibom_et_al_2013}.  The reason is that
during the main Kepler mission (2009--2013), only four open
clusters were known in the Kepler field: NGC\,6866, NGC\,6811,
NGC\,6819, and NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.  Since that time, analyses of the
kinematic, photometric, and astrometric Gaia data have expanded our
knowledge of open cluster and moving group memberships \citep[{\it
e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,Kerr2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Appendix~\ref{app:targetlist}).

@@

A previous clustering analysis of the Gaia data reported that
Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) is a member of the
\cn\ \citep{kounkel_untangling_2019}.  Given the previous statistical
validation of the close-in mini-Neptune Kepler 1627b
\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
we examined the age of the star and its putative host cluster more
closely (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr
cluster is \clusterage\ old, and that \sn\ is indeed a member of the
cluster.  Focusing on the planet itself, we confirm that despite the
presence of a previously unreported binary companion, hereafter
Kepler\,1627B, the planet orbits the G-dwarf primary on an orbit that
is likely aligned with the stellar rotation axis
(Section~\ref{sec:planet}).  We conclude by highlighting broader
implications for our ability to age-date a statistical sample of
planets (Section~\ref{sec:conc}).

@@

In the context of the transiting planet population, \pn\ is among the
youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
with precise ages below 100\,Myr include K2-33, DS~Tuc, HIP~67522,
TOI~837, the two-planet AU~Mic system, and the four-planet V1298~Tau
system
\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020,martioli_aumicbc_2021}.
\pn\ is one of the smallest planets in this sample ($3.34\pm
0.18\,R_\oplus$), which could be linked to the selection effects
imposed by the spot-induced photometric variability at very young ages
\citep[{\it e.g.},][]{zhou_2021_tois}.  Assuming that these young
planets have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$,
their large sizes and presumed low densities would be in accord with
the expectation that mini-Neptunes start their lives with large
primordial atmospheres that are shed over the first gigayear
\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  Our
suggestion of an aligned orbit is consistent with a quiescent
disk-driven migration history, and differs from misalignments that
have been observed for some older
mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
Spectroscopic follow-up observations of the transit remain necessary
however to verify whether the TTV-slope correlation is yielding the
correct orbital geometry.

@@

The catalogs we included for which ages were not immediately available
were those of \citet{CottenSong2016}, \citet{Oh2017},
\citet{Zari2018}, \citet{Gagne2018b},
\citet{Gagne2018a}, \citet{Gagne2018c}, and \citet{Ujjwal2020}.  While
in principle the moving group members discussed by
\citet{Gagne2018b,Gagne2018a,Gagne2018c} and \citet{Ujjwal2020} have
easily associated ages, our SIMBAD cross-match did not retain the moving group
identifiers given by those studies, which should therefore be recovered
tools such as BANYAN
$\Sigma$.\footnote{\url{http://www.exoplanetes.umontreal.ca/banyan/banyansigma.php}}.
We also included the SIMBAD object identifiers $\texttt{TT*}$,
$\texttt{Y*O}, $\texttt{Y*?}, $\texttt{TT?}$, and $\texttt{pMS*}$.
Finally, we  included every star in the NASA Exoplanet Archive
planetary system ($\texttt{ps}$) table that had a Gaia identifier available
\citep{NASAExoArchive_ps_20210506}.  If the age had finite
uncertainties, we also included it, since stellar ages determined
through the combination of isochrone-fitting and transit-derived
stellar densities typically have higher precision than from isochrones
alone.

@@

Figure~\ref{fig:XYZvtang} shows stars reported by
\citet{kounkel_untangling_2019} to be in the group ``Theia 73'', which
was cross-matched by Kounkel \& Covey as being ``Stephenson 1''.
\citet{kounkel_untangling_2019} reported \noriginal\ stars to be
present in this cluster.  For our Figure, galactic positions were
calculated and plotted only for stars with parallax signal-to-noise
exceeding 20.  The location of the Sun is shown on the plots.  The
smattering of reported cluster members (gray and black points) has a
significantly different structure that the cluster initially
identified by \citet{stephenson_possible_1959} and corroborated by
\citet{eggen_photometric_1968}.  While the non-uniform ``clumps''
might be part of a real structure, they could also be an artifact of
the data processing steps performed by
\citet{kounkel_untangling_2019}.  We therefore opted to only consider
stars in the immediate kinematic and spatial vicinity of \sn.  The
tangential velocities relative to \sn\ are shown in the bottom right
panel.  These are computed by assuming that every star has the same
three-dimensional spatial velocity as \sn, where we assume a systemic
radial velocity of $-16.7 \pm 0.2$\,\kms\ based on the HIRES
spectra.  The
relevant projection effects are then taken into account, as discussed
by {\it e.g.}, \citet{Meingast2021} and \citet{bouma_2021_ngc2516}.
We performed the actual selection by then manually drawing lassos with
the interactive \texttt{glue} visualization tool
\citep{beaumont_2014_13866} in the four projections shown in
Figure~\ref{fig:XYZvtang}.  While other cluster members likely exist
outside of our selection region (and our selection also includes some
field star interlopers), our aim is to verify the existence of the
cluster in the vicinity of Kepler\,1627, and to measure its age.  The
procedure we have adopted enables both tasks.

@@

The presence of this additional scatter prevents a clear detection of
the RM effect.  The reason can be understood through a
model-comparison analysis.  If we compare the model with a quadratic trend and
the RM effect (Figure~\ref{fig:obliquity}) against a model with a
linear trend and the RM effect, or even a model with no RM effect at
all, then the respective Bayesian Information Criterion (BIC) values
are as follows.
% from logs/rm_freejitter.log
% \begin{align}
%   \chi^2_{\rm red} &= 1.45 \quad ({\rm Quadratic+RM}) \nonumber \\
%   \chi^2_{\rm red} &= 1.37 \quad ({\rm Linear+RM}) \nonumber \\
%   \chi^2_{\rm red} &= 1.22 \quad ({\rm Only\ Quadratic}),
% \end{align}
% and
\begin{align}
  {\rm BIC} &= 251.7  \quad ({\rm Quadratic+RM}) \nonumber \\
  {\rm BIC} &= 257.7  \quad ({\rm Linear+RM}) \nonumber \\
  {\rm BIC} &= 242.8  \quad ({\rm Only\ Quadratic}).
\end{align}
There is therefore no evidence to prefer the model with the RM effect
against a model that only accounts for the stellar variability.
The ``only quadratic'' model does particularly well because it can
inflate the jitter term to account for any additional scatter during
the transit, and it has fewer free parameters.  However, we cannot
justify a physical prior on the jitter term, because we do not
understand the origin of the exposure-to-exposure scatter.  As noted
above, the velocity deviations from starspots are expected to have
contributions at the stellar rotation frequency, or harmonics thereof.
This jitter is present on the exposure timescale (15 minutes), which
is only 0.4\% of the stellar rotation period; it is not obvious that
starspots could be the culprit.

>>>>

% 208 words (250 max)
\begin{abstract}
  Kepler\,1627A is a G8V star previously known to host a
  $3.3\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  observed by the Kepler space telescope because it is nearby
  ($d=329\pm3\,{\rm pc}$) and it resembles the Sun.  Here we show
  using Gaia kinematics, TESS stellar rotation periods, and
  spectroscopic lithium abundances that Kepler\,1627 is a member of
  the \clusterage\ old $\delta$~Lyr cluster.  To our knowledge,
  Kepler\,1627Ab is the youngest planet with a precise age yet found
  by the main Kepler mission.  From unresolved starspot-crossing
  events in the Kepler photometry and a Keck/HIRES transit we find
  that the planet is likely on a prograde orbit.  We also briefly
  explore the flare arrival time distribution in the short cadence
  Kepler photometry.  The $\delta$~Lyr cluster is one of many stellar
  groups whose properties have been clarified using Gaia; many other
  exoplanet hosts belong to these new clusters.  Verifying the results
  of Gaia clustering analyses using TESS rotation periods offers a
  promising path toward expanding the census of age-dated planets, and
  our corresponding understanding of planetary evolution.
\end{abstract}

@@

The K2 and TESS missions have now enabled the detection of about ten
close-in planets younger than 100 million years, all smaller than
Jupiter
\citep{Mann_K2_33b_2016,David_et_al_2017,david_four_2019,newton_tess_2019,bouma_cluster_2020,plavchan_planet_2020,rizzuto_tess_2020,martioli_aumicbc_2021}.
The Kepler mission however has not yielded any planets with precise
ages below one gigayear \citep{Meibom_et_al_2013}.  The reason is that
during the main Kepler mission (2009--2013), only four open clusters
were known in the Kepler field: NGC\,6866, NGC\,6811, NGC\,6819, and
NGC\,6791, with ages spanning 0.7\,Gyr to 9\,Gyr
\citep{meibom_kepler_2011}.  Though isochronal and gyrochronal
analyses suggest that younger Kepler planets do exist
\citep{berger_identifying_2018,david_sizes_2021}, accurate and precise
age measurements typically require an ensemble of stars.  Fortunately,
recent analyses of the kinematic, photometric, and astrometric Gaia
data have significantly expanded our knowledge of cluster memberships
\citep[{\it
e.g.},][]{cantatgaudin_gaia_2018,zari_3d_2018,kounkel_untangling_2019,Meingast2021,Kerr2021}.
As part of our Cluster Difference Imaging Photometric Survey (CDIPS,
\citealt{bouma_cdipsI_2019}), we concatenated the available analyses
from the literature, which yielded a list of candidate young and
age-dated stars (see Appendix~\ref{app:targetlist}).

@@

A previous clustering analysis of the Gaia data reported that
Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) is a member of the
\cn\ \citep{kounkel_untangling_2019}.  Given the previous statistical
validation of the close-in mini-Neptune Kepler 1627b
\citep{2012ApJS..199...24T,morton_false_2016,thompson_planetary_2018},
we examined the age of the star and its putative host cluster more
closely (Section~\ref{sec:cluster}).  We find that the $\delta$ Lyr
cluster is \clusterage\ old, and that \sn\ is indeed a member of the
cluster.  Focusing on the planet itself, we confirm that despite the
presence of a previously unreported binary companion, hereafter
Kepler\,1627B, the planet orbits the G-dwarf primary on an orbit that
is likely prograde
(Section~\ref{sec:planet}).  We conclude by highlighting broader
implications for our ability to age-date a statistical sample of
planets (Section~\ref{sec:conc}).

@@

In the context of the transiting planet population, \pn\ is among the
youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
with precise ages below 100\,Myr include K2-33, DS~Tuc, HIP~67522,
TOI~837, the two-planet AU~Mic system, and the four-planet V1298~Tau
system
\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020,martioli_aumicbc_2021}.
\pn\ is one of the smallest planets in this sample ($3.34\pm
0.18\,R_\oplus$), which could be linked to the selection effects
imposed by the spot-induced photometric variability at very young ages
\citep[{\it e.g.},][]{zhou_2021_tois}.  Assuming that these young
planets have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$,
their large sizes and presumed low densities would be in accord with
the expectation that mini-Neptunes start their lives with large
primordial atmospheres that are shed over the first gigayear
\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  Our
suggestion of a prograde is consistent with a quiescent
disk-driven migration history, and differs from misalignments that
have been observed for some older
mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
Spectroscopic follow-up observations of the transit remain necessary
however to verify whether the TTV-slope correlation is yielding the
correct orbital geometry.

@@

The catalogs we included for which ages were not immediately available
were those of \citet{CottenSong2016}, \citet{Oh2017},
\citet{Zari2018}, \citet{Gagne2018b}, \citet{Gagne2018a},
\citet{Gagne2018c}, and \citet{Ujjwal2020}.  While in principle the
moving group members discussed by
\citet{Gagne2018b,Gagne2018a,Gagne2018c} and \citet{Ujjwal2020} have
easily associated ages, our SIMBAD cross-match did not retain the
moving group identifiers given by those studies, which should
therefore be recovered using tools such as BANYAN
$\Sigma$.\footnote{\url{http://www.exoplanetes.umontreal.ca/banyan/banyansigma.php}}.
We also included the SIMBAD object identifiers $\texttt{TT*}$,
$\texttt{Y*O}, $\texttt{Y*?}, $\texttt{TT?}$, and $\texttt{pMS*}$.
Finally, we  included every star in the NASA Exoplanet Archive
planetary system ($\texttt{ps}$) table that had a Gaia identifier
available \citep{NASAExoArchive_ps_20210506}.  If the age had finite
uncertainties, we also included it, since stellar ages determined
through the combination of isochrone-fitting and transit-derived
stellar densities typically have higher precision than from isochrones
alone.

@@

Figure~\ref{fig:XYZvtang} shows stars reported by
\citet{kounkel_untangling_2019} to be in the group ``Theia 73'', which
was cross-matched by Kounkel \& Covey as being ``Stephenson 1''.
\citet{kounkel_untangling_2019} reported \noriginal\ stars to be
present in this cluster.  For our Figure, galactic positions were
calculated and plotted only for stars with parallax signal-to-noise
exceeding 20.  The location of the Sun is shown on the plots.  The
smattering of reported cluster members (gray and black points) has a
significantly different structure than the cluster initially
identified by \citet{stephenson_possible_1959} and corroborated by
\citet{eggen_photometric_1968}.  While the non-uniform ``clumps''
might be part of a real structure, they could also be an artifact of
the data processing steps performed by
\citet{kounkel_untangling_2019}.  We therefore opted to only consider
stars in the immediate kinematic and spatial vicinity of \sn.  The
tangential velocities relative to \sn\ are shown in the bottom right
panel.  These are computed by assuming that every star has the same
three-dimensional spatial velocity as \sn, where we assume a systemic
radial velocity of $-16.7 \pm 0.2$\,\kms\ based on the HIRES
spectra.  The
relevant projection effects are then taken into account, as discussed
by {\it e.g.}, \citet{Meingast2021} and \citet{bouma_2021_ngc2516}.
We performed the actual selection by then manually drawing lassos with
the interactive \texttt{glue} visualization tool
\citep{beaumont_2014_13866} in the four projections shown in
Figure~\ref{fig:XYZvtang}.  While other cluster members likely exist
outside of our selection region (and our selection also includes some
field star interlopers), our aim is to verify the existence of the
cluster in the vicinity of Kepler\,1627, and to measure its age.  The
procedure we have adopted enables both tasks.

@@

The presence of this additional scatter prevents a clear detection of
the RM effect.  The reason can be understood through a
model-comparison analysis.  If we compare the model with a quadratic trend and
the RM effect (Figure~\ref{fig:obliquity}) against a model with a
linear trend and the RM effect, or even a model with no RM effect at
all, then the respective Bayesian Information Criterion (BIC) values
are as follows.
% from logs/rm_freejitter_FINAL.log
\begin{align}
  {\rm BIC} &= 251.7  \quad ({\rm Quadratic+RM}) \nonumber \\
  {\rm BIC} &= 257.7  \quad ({\rm Linear+RM}) \nonumber \\
  {\rm BIC} &= 247.1  \quad ({\rm Only\ Quadratic}).
\end{align}
There is therefore no evidence to prefer the model with the RM effect
against a model that only accounts for the stellar variability.
The ``only quadratic'' model does particularly well because it can
inflate the jitter term to account for any additional scatter during
the transit, and it has fewer free parameters.  However, we cannot
justify a physical prior on the jitter term, because we do not
understand the origin of the exposure-to-exposure scatter.  As noted
above, the velocity deviations from starspots are expected to have
contributions at the stellar rotation frequency, or harmonics thereof.
This jitter is present on the exposure timescale (15 minutes), which
is only 0.4\% of the stellar rotation period; it is not obvious that
starspots could be the culprit.
